PLAE enables quick and easy searching across one million ocular transcriptomes

### 2022-03-22 manuscript_00.Rmd
text / code tweaks

library(pool)
library(RSQLite)
meta_filter_v0 <- fst::read_fst("/Volumes/McGaughey_S/scEiaD/paper_data/2021_03_17_meta_filter.fst") %>% 
  mutate(CellType_predict = case_when(CellType_predict == 'AC/HC_Precurs' ~ 'AC/HC Precursor',
                                      TRUE ~ CellType_predict),
         CellType= case_when(CellType == 'AC/HC_Precurs' ~ 'AC/HC Precursor',
                                      TRUE ~ CellType))
# remove eric bharti 2D/3D RPE
meta_filter <- fst::read_fst("/Volumes/McGaughey_S/scEiaD_v3/2021_11_09_meta_filter.fst") %>% filter(study_accession != 'Bharti_Nguyen_iRPE_2D_3D') %>% 
  mutate(CellType_predict = case_when(CellType_predict == 'AC/HC_Precursor' ~ 'AC/HC Precursor',
                                      TRUE ~ CellType_predict),
         CellType= case_when(CellType == 'AC/HC_Precursor' ~ 'AC/HC Precursor',
                                      TRUE ~ CellType))

@@

# Figure 1 Distribution of Tissues and Cell Types in scEiaD
```{r, fig.width=8, fig.height=5}
# organ
meta_filter <- meta_filter %>% mutate(organism = factor(organism, levels = meta_filter$organism %>% unique()))
a <- meta_filter %>%
  mutate(Organ = case_when(Organ == 'Eye' ~ 'Eye',
                           TRUE ~ 'Body'),
         Tissue = gsub('\\.|_', ' ', Tissue),
         Tissue = gsub("RPE$","RPE-Choroid", Tissue),
         Tissue = gsub(" and Aorta","", Tissue)) %>% 
  filter(Organ == 'Body') %>% 
  filter(Tissue != 'Brain Choroid Plexus') %>% 
  group_by(Organ, Tissue,organism, study_accession) %>%
  summarise(Count = n()) %>%
  ungroup() %>% 
  mutate(organism = factor(organism, levels = meta_filter$organism %>% unique())) %>% 
  ggplot(aes(x=Tissue,y=Count, fill = organism, group = organism)) +
  ylab('Number of Cells') +
  geom_bar(stat='identity', position = 'dodge') +
  cowplot::theme_cowplot() +
  coord_flip() +
  # scale_y_continuous(breaks = c(2500, 5000, 7500, 10000, 12500, 84000)) +
  # ggbreak::scale_y_break(c(13000, 82000)) +
  scale_fill_discrete(drop=TRUE,
                      limits = levels(meta_filter$organism)) +
  theme(legend.position = "none") +
  scale_y_continuous(expand = c(0, 0)) + 
  ylab('') + xlab('')

@@

a_aa_b <- plot_grid(a,aa,b, ncol = 1, align = 'hv', rel_heights = c(10, 1.5,3.5), labels = 'a')
c_d <- plot_grid(c,d, rel_heights = c(2,10), ncol = 1, align= 'hv', axis = 'l',labels = c('b','c'))
plot_grid(a_aa_b, c_d, ncol = 2, rel_widths = c(1,1.2))
```

@@

```{r, fig.width=10, fig.height=12}
source('src/make_exp_plot.R')
scEiaD_2020_v00 <- pool::dbPool(RSQLite::SQLite(), dbname =  '~/data/scEiaD/MOARTABLES__anthology_limmaFALSE___5000-transform-counts-universe-batch-scVIprojectionSO-8-0.1-50-5.sqlite')
scEiaD_2020_v01 <- pool::dbPool(RSQLite::SQLite(), dbname =  '/Volumes/McGaughey_S/scEiaD_v3/MOARTABLES__anthology_limmaFALSE___6000-counts-universe-batch-scVIprojection-10-15-0.1-50-20.sqlite')

@@

## Differential testing identifies many genes consistently enriched in various ocular cell types

## Ocular specific gene signatures

## Fovea vs peripheral photoreceptor validation

@@

## Design of scEiaD allows for projection of outside data

>>>>

library(pool)
library(RSQLite)
meta_filter_v0 <- fst::read_fst("~/data/scEiaD/paper_data/2021_03_17_meta_filter.fst") %>% 
  mutate(CellType_predict = case_when(CellType_predict == 'AC/HC_Precurs' ~ 'AC/HC Precursor',
                                      TRUE ~ CellType_predict),
         CellType= case_when(CellType == 'AC/HC_Precurs' ~ 'AC/HC Precursor',
                             TRUE ~ CellType))
# remove eric bharti 2D/3D RPE
meta_filter <- fst::read_fst("~/data/scEiaD_2022_02/2022_03_22_meta_filter.fst") %>% filter(study_accession != 'Bharti_Nguyen_iRPE_2D_3D') %>% 
  mutate(CellType_predict = case_when(CellType_predict == 'AC/HC_Precursor' ~ 'AC/HC Precursor',
                                      TRUE ~ CellType_predict),
         CellType= case_when(CellType == 'AC/HC_Precursor' ~ 'AC/HC Precursor',
                             TRUE ~ CellType),  
         Tissue = case_when(grepl('Iris', Tissue) ~ 'Iris',
                            TRUE ~ Tissue))

@@

# Figure 1 Distribution of Tissues and Cell Types in scEiaD
```{r, fig.width=5.5, fig.height=5}
# organ
meta_filter <- meta_filter %>% mutate(organism = factor(organism, levels = meta_filter$organism %>% unique()))
a <- meta_filter %>%
  mutate(Organ = case_when(Organ == 'Eye' ~ 'Eye',
                           TRUE ~ 'Body'),
         Tissue = gsub('\\.|_', ' ', Tissue),
         Tissue = gsub("RPE$","RPE-Choroid", Tissue),
         Tissue = gsub(" and Aorta","", Tissue)) %>% 
  filter(Organ == 'Body') %>% 
  filter(Tissue != 'Brain Choroid Plexus') %>% 
  group_by(Organ, Tissue,organism, study_accession) %>%
  summarise(Count = n()) %>%
  ungroup() %>% 
  mutate(organism = factor(organism, levels = meta_filter$organism %>% unique())) %>% 
  ggplot(aes(x=Tissue,y=Count, fill = organism, group = organism)) +
  ylab('Number of Cells') +
  geom_bar(stat='identity', position = 'dodge') +
  cowplot::theme_cowplot() +
  coord_flip() +
  # scale_y_continuous(breaks = c(2500, 5000, 7500, 10000, 12500, 84000)) +
  # ggbreak::scale_y_break(c(13000, 82000)) +
  scale_fill_discrete(drop=TRUE,
                      limits = levels(meta_filter$organism)) +
  theme(legend.position = "none") +
  scale_y_continuous(expand = c(0, 0)) + 
  ylab('') + xlab('')

@@

a_aa_b <- plot_grid(a,aa,b, ncol = 1, align = 'hv', rel_heights = c(10, 1.5,3.5), labels = 'a')
c_d <- plot_grid(c,d, rel_heights = c(2,10), ncol = 1, align= 'hv', axis = 'l',labels = c('b','c'))
plot_grid(a_aa_b, c_d, ncol = 2, rel_widths = c(1,1.6))
```

@@

```{r, fig.width=10, fig.height=12}
source('src/make_exp_plot.R')
scEiaD_2020_v00 <- pool::dbPool(RSQLite::SQLite(), dbname =  '~/data/scEiaD/MOARTABLES__anthology_limmaFALSE___5000-transform-counts-universe-batch-scVIprojectionSO-8-0.1-50-5.sqlite')
scEiaD_2020_v01 <- pool::dbPool(RSQLite::SQLite(), dbname =  '~/data/scEiaD_2022_02/MOARTABLES__anthology_limmaFALSE___4000-counts-universe-study_accession-scANVIprojection-15-5-0.1-50-20.sqlite')

@@

## Differential testing identifies many genes consistently enriched in various ocular cell types 
```{r}
scEiaD_2020_v01 %>% tbl('wilcox_diff_testing') %>% filter(Group == 'CellType (Predict)', FDR < 1e-5, mean_auc > 0.7 ) %>% group_by(Base) %>% summarise(Count = n()) %>% data.frame()
```

## Fovea vs peripheral photoreceptor multi-study validation

@@

## Projection of organoid data onto scEiaD demonstrates [SOMETHING]
